\section*{Acknowledgements}
\label{sec:acknowledgements}
\addcontentsline{toc}{chapter}{\nameref{sec:acknowledgements}}

This thesis was made possible with a support of many different people and
resources. First and foremost is the support of my research supervisor Prof.
Nikolas Provatas whose drive and enthusiasm for his research is paralleled only
by his support and guidance in his teaching. He gave me a world of questions
I'll never tire of considering, something I can never thank him enough for.  

I would also like to thank Prof. David Ronis, for many great discussions
especially in the murky waters of non-equilibrium statistical mechanics.  His
course notes and references proved invaluable in the writting of this thesis
(the appendix on generialized Einstein relations most especially).  About 7
years prior to the writing of this thesis, he entertained my question "How do
you know if you've chosen the correct reaction co-ordinate?" in his
Introduction to Physical Chemistry course. It seems in retrospect that I
haven't stopped asking that question.

Thank you Camille Bowness for being so supportive throughout this entire
degree, for many discussions over morning espresso or on the walk to campus.
Thanks as well for your help in translating the abstract.

I swarm of great people helped in the editting of this thesis including Quentin
Stoyel, Motoki Nakamura, Paul Jreidini and, of course, Nikolas Provatas. Thank
you all for your time. I can't thank you enough for you patience and advice.

A huge thank you to my family and friends for supporting me while I took these
years to take a magnifying glass to the world. I couldn't have done it without
you.

Finally, this research was made possible with resources from Compute Canada and
Calcul Quebec and with funding from the Canada Research Chairs Program.

