\label{chapter:introduction}

% Thesis introduction outline: https://student.unsw.edu.au/introductions

% 1) State the general topic and give some background

  
\nc{ The study of alloys in materials physics is a pursuit of incredibly broad impact since 
the functional properties of materials depend on their microstructure, which forms through 
non-equilibrium phase transformations during the process of forming a material. As a 
result, industries  as diverse  as those dealing in commercial products based on steel 
and aluminium to the burgeoning  markets dealing in nano-fabrication and 
optoelectronics are affected by research in alloy materials.  One of the most useful 
paradigms for understanding complex alloy microstructure is that of the binary alloy. }

One surprising aspect of binary alloys is the rich diversity of properties and
behaviours they display. Because material properties depend on the
microstructural details of the material, they have a strong processing path dependence.
Grain boundaries, vacancies, dislocations and other microstructural artifacts
are all intimately tied to the manufacturing process of the alloy. This means
that the study of solids can never be completely separated from the study of
solidification. As such, the diversity of material properties and behaviours we
see in binary alloys can be directly attributed to the diversity of processes
for their construction.

Given the importance of binary systems, it is critical to construct models that
can explain the diversity of behaviour we see in them. At the moment, models of
alloy solidification can be categorized by the length and time scales they accurately
describe. At macroscopic length and time scales, we have continuum methods of
heat and mass transport and associated finite element methods of analysis. These
methods are appropriate for studying large castings, for example. On length
scales $\mathcal{O}(10^{-6}m)$ to $\mathcal{O}(10^{-3}m)$ we use
\textit{Phase Field} methods to study phenomena such as dendritic growth and
chemical segregation. On still finer length scales from $\mathcal{O}(10^{-9}m)$
to $\mathcal{O}(10^{-6}m)$ and on relatively long timescales, we have the
methods of \textit{Phase Field Crystal} (PFC) theory and dislocation dynamics.
These methods are appropriate for studying nanoscopic changes that occur on
diffusive timescales such as dislocation motion, creep, grain boundary motion
and micro segregation. At a still finer scale and on very short time scales
($\mathcal{O}(10^{-12}s)$) we have the methods of molecular dynamics and density
functional theory. These methods are appropriate for the study of transport
coefficients and interaction potentials.

% 2) Outline the current situation

\nc{ In this thesis, we'll focus on extending a branch of binary PFC theory known as the binary 
{\it XPFC} model --where the "X" in XPFC signifies a class of PFC models  constructed 
to controllably simulate a robust range of metallic and non-metallic crystal 
symmetries compared to the original PFC models.}  PFC binary models have been 
successful in describing a broad selection of phenomena in binary alloys.
These successes include \nc{eutectic and dendritic solidification \cite{ELDER07}},
 the Kirkendall effect \cite{ELDER11_KIRKENDALL, LU15},
solute drag \cite{GREENWOOD12}, clustering and precipitation \cite{FALLAH12,
FALLAH13, FALLAH13_AlCu_experiment}, colloidal ordering in drying suspensions
\cite{GANAI13}, epitaxial growth and island formation \cite{ELDER10_NANOISLAND,
LU16}, and ordered crystals \cite{ALSTER17} to name a few. 

The PFC theory is derived from Classical Density Functional Theory (CDFT) and as
such, it can be considered a simplified density functional theory. 
In practice, two different variants of the PFC theory are used, as alluded to above: the 
original model developed by Elder \textit{et al} \cite{ELDER07} and the Structural Phase 
Field Crystal (XPFC) model developed by Greenwood \textit{et al.} \cite{GREENWOOD11_BINARY}.

\nc{The original model was the first PFC theory of binary alloys and contains some important 
 physical properties of binary alloys.} However, it is a very reduced form of CDFT 
and it therefore lacks completeness in its ability to describe binary alloys. Specifically, the original 
model uses an expansion in concentration {\nc{that is actually a density difference not a concentration}, 
and the model has a limited ability to describe a realistic or robust range of 
phase diagrams. The original model also uses a very simplified correlation kernel which limits its 
ability to describe a variety of crystal lattice structures.

The XPFC model is an improvement that ameliorates the above problems. The
concentration is left unexpanded allowing for construction of realistic global
phase diagrams instead of local expansions. More significantly, the XPFC model provides a
phenomenology for modelling \nc{two-point correlation functions} that succeeded in describing
solidification of a variety of lattice structures, \nc{as well as transformations between different crystal 
lattices.}  Simplifications of the multi-modal approach first introduce with the XPFC formalism has been used to
produce hexagonal, square, kagome, honeycomb, rectangular and other lattices in
2 dimensions \cite{MKHONTA13}.

% 3) Evaluate the current situation and identify a gap

In introducing its phenomenology for modelling correlation functions, the \nc{binary} XFPC
theory tacitly assumes that there is some preferred structure at high
concentration and some other structure preferred at low concentration. This
assumption can be limiting in situations that have a specific crystalline
structure at intermediate concentrations, such as materials with a syntectic
phase diagram. At the syntectic point a solid of intermediate concentration
solidifies along the interface between a solute rich and solute poor liquid. The
XPFC model also assumes no long wavelength correlations in the concentration
field which, in practice, means the model has an ideal free energy of mixing.
This is another limitation of the XPFC model because the enthalpy of mixing is
not generally zero for alloy systems.

% 5) State the research goals and aims

\nc{The goal of the current research is threefold: The first two goals are
to present two important improvements to the binary XPFC theory.} The first 
improvement is a more general phenomenology for modelling pair correlation 
functions of a binary material. The second improvement is to extend the free energy 
of mixing beyond ideality to account for circumstances when the heat of mixing 
is not negligible. \nc{The third goal is to use the new XPFC model derived herein 
to the elucidate the multi-step nucleation process in an experimentally relevant 
nanoparticles system} \nb{(state system and add reference to Singapore group)}

% 6) Outline the order of information

The remainder of this thesis is divided into 5 chapters:
%
\begin{description}
    \item [Chapter \ref{chapter:cdft_intro}] { Classical Density Functional
        Theory (CDFT) is introduced and derived from fundamental principles of
        quantum statistical mechanics.
    }
    \item [Chapter \ref{chapter:cdft_of_freezing}] { CDFT theory of
        solidification is described and discussed. The density functional
        theory is extended to a dynamic, non-equilibrium theory, and the Phase
        Field Crystal (PFC) Theory \nc{is introduced from it}  as a simplified density
        functional theory.
    }
    \item [Chapter \ref{chapter:binary}] {Binary PFC theory is \nc{derived and
        previous simplified alloy PFC models} are summarized and discussed.
    }
    \item [Chapter \ref{chapter:improvements}] {Improvements to the XPFC binary alloy theory 
        are \nc{derived}. This chapter contains novel contributions to the field.
    }
    \item [Chapter \ref{chapter:applications}] {
    \nc{The new XPFC alloy model derived herein is applied to} model to the problem of multi-step nucleation
        of nanoparticles in the system \nb{state system,} and potential future applications of this model are discussed.
    } 
\end{description}
%
